Journey: A Study of Studying Large Memory Systems (TODO: should we get a new title?)

### 2017-04-20 paper.tex
Worked on background section

(TODO: do we need to refocus this section? or move parts of it to an appendix?)

@@

@@

@@

@@

@@

(TODO: optional: trying pulling out a DIMM??)

>>>>

@@

Under memory pressure, the kernel may choose to write some pages of memory to
disk and then free those pages for use. Memory mapped files are written back to
their backing files, but anonymous memory is written to portion of the disk
called swap space. We focus on anonymous memory in our experiments.

Before swapping a page out to disk, the kernel must reserve a slot for the page
in the swap space, unmap it from the process's address space, and record which
swap slot corresponds to the page.

On Linux, the system may have multiple swap spaces in the form of a swap disk
partition or of swapfiles on a filesystem. Both have a similar format on disk,
in which the space is broken into a number of swap slots large enough for one
page. For each of these swap spaces, the kernel keeps a swap map, which records
the usage of each swap slot so that it can find and allocate them quickly.

Unmapping a page from a process's address space involves simply marking its page
table entry ``Not Present''. Rather than using a separate data structure, Linux
uses the unused page table entry to record information about which swap space
and slot is used to swap out the page.

@@

To determine which pages to swap out, Linux uses the Page Frame Reclamation
subsystem. Choosing which pages to swap out is important because if the system
swaps out pages that are about to be used, thrashing may be induced, in which
the system wastes lots of time swapping out a page just swap it back in.

The Linux Page Frame Reclamation Algorithm (PFRA) attempts to approximate LRU
order. It uses a clock-like algorithm, in which a list of pages is scanned
twice. Roughly speaking, if a page is not used between scans, then it is a
condidate for reclamation.

The PFRA runs primarily as a kernel thread called kswapd. kswapd runs at regular
intervals and sleeps in between. Also, it may be invoked if the system detects
that memory has gone below a threshold. kswapd will continue to run until enough
pages are freed to bring the free memory pool above some threashold. Then,
kswapd goes back to sleep.  One concern with large memory systems is that kswapd
may expend significant CPU time scanning through hundreds of thousands of pages.

@@

\subsection{\texttt{vm\_area\_struct}}

@@

\subsection{Swapping Memory Overhead}

@@

\subsection{kswapd}
